---
title: "covid 19"
author: "Juan Moreno"
date: "7/21/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

First we need a basic table to start out with as this can be used to make charts with later on.  

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
BP <- kRdata %>% group_by(result, sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# frequency plot

basicC <- ggplot(BP,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

# percent plot

basicCC <- ggplot(BP,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In the graphs above you can see frequency wise it appears more male die then female, then looking at the percent graph we can see that males die more then females.  
  
## Numerical results


## Inferential results 




## Conclusions

