---
title: "covid 19"
author: "Juan Moreno"
date: "7/21/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

First we need a basic table to start out with as this can be used to make charts with later on.  

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
BP <- kRdata %>% group_by(result, sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# frequency plot

basicC <- ggplot(BP,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

# percent plot

basicCC <- ggplot(BP,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In the graphs above you can see frequency wise it appears more male die then female, then looking at the percent graph we can see that males die more then females.  
  
## Numerical results

```{r}
table2 <- xtabs(~result + sex, data=kRdata)
rowPerc(table2)
colPerc(table2)
```

Now looking at the numerical results we percentages both   in columns and in rows that was shown in the bar graphs   above. That was not mentioned or explained well in the   earlier part where we had a simple  data chart  

## Inferential results 

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Take notice of the P-value from the fisher test, notice   how small the p-value is, this means that there is a   difference of some kind and its one that is backed up   with chi-sq test.  


## Conclusions

From looking at all the data we can see that being a male and catching covid will increase the likely hood of dying then if a person was a female.
